---
title: "Data Exploration and Dynamics"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: TRUE
    toc_depth: 4            # Define the depth of the TOC (increase as needed)
    toc_float:
      collapsed: TRUE      # Whether the TOC starts collapsed
      smooth_scroll: TRUE   # Smooth scrolling when a TOC entry is clicked
fontsize: 12pt              # Set font size
linkcolor: blue             # Hyperlink color (for PDF output)


vignette: >
  %\VignetteIndexEntry{Data Exploration and Dynamics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup_general, include = FALSE}
knitr::opts_chunk$set(
   echo = TRUE,
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```




```{r setup_lib, include=FALSE}
if (!requireNamespace("htmltools", quietly = TRUE)) {
  stop("Package 'htmltools' is required to run this function. Please install it.")
}
if (!requireNamespace("magick", quietly = TRUE)) {
  stop("Package 'magick' is required to run this function. Please install it.")
}
library(growthTrendR)
library(data.table)
library(ggplot2)
library(sf)


```

This vignette presents two analyses of growth-related attributes: the Frequency Report summarizes sample distributions across spatial locations, while the Spatial–Temporal Report visualizes interpolated attributes over space and time.

<br>



```{r data_qa, echo= FALSE, eval = FALSE, eval=FALSE}

```


## Frequency

The Frequency Report provides frequency distributions across locations, summarizing patterns at the site, tree, and sample levels, and highlighting data completeness, variability, and potential gaps for further analysis.

```{r freq}
# loading processed data
# otherwise need to run CFS_format() first as done in data report
 dt.samples_trt <- readRDS(system.file("extdata", "dt.samples_trt.rds", package = "growthTrendR"))
 # Compute frequency statistics at radius level
 dt.freq <- CFS_freq(
   data = dt.samples_trt$tr_all_wide,
   freq.label_data = "demo-samples",
   freq.uid_level  = "uid_radius"
 )
```


#### **arguments of CFS_freq() function:**

**data**

Metadata table generated by the function CFS_format().

**freq.label_data**

A short description of the dataset used. This text will appear in the report as the data source for the generated figures.

**freq.uid_level**

Specifies the hierarchical level at which frequencies are counted (e.g., site, tree, sample, or radius).



### **generate report:**

```{r freq_report_demo, eval= FALSE}
generate_report(robj = dt.freq)
```



<br>

#### **arguments of the generate_report() function:**

**robj **

The input for the frequency report is the output of the CFS_freq() function, which assigns the class "cfs_freq" to the resulting object.

**output_file **

The output_file argument allows users to generate an HTML report at a specified path and file name. If set to NULL (default), the report opens automatically in the browser.

```{r freq_report, include = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
outfile_freq <- tempfile(fileext = ".html")
generate_report(robj = dt.freq, output_file = outfile_freq)
```

```{r freq_inclu, echo = FALSE, results = 'asis'}

cat('
<div style="margin: 20px 0; padding: 10px; 
            box-shadow: 0 2px 8px rgba(0,0,0,0.1); 
            border-radius: 8px;">
')

htmltools::includeHTML(outfile_freq)  # embed in vignette
cat('</div>')
```
------------------------------------------------------------------------


## Spatial-Temporal report

This report presents the spatial–temporal dynamics of growth-related attributes (e.g., ring width, basal area increment, chronology indices, or model-predicted values). The variables are interpolated across the study region using the CFS_mapping() function with the inverse distance weighting (IDW) method, and the resulting raster layers are compiled into a GIF for visualization.


```{r st, message = FALSE, warning = FALSE, results = 'hide'}

# loading processed data
dt.samples_trt <- readRDS(system.file("extdata", "dt.samples_trt.rds", package = "growthTrendR"))


# data processing

cols.meta = c("uid_tree", "uid_site", "longitude", "latitude", "species")

dt.mapping <- dt.samples_trt$tr_all_wide[
  , c(..cols.meta, as.character(1991:1995)), with = FALSE]

results_mapping <- CFS_mapping(data = dt.mapping, year.span = c(1991,1995))


```

#### **arguments of CFS_mapping() function:**

**data**

The input dataset must include at least five columns: uid_tree, uid_site, longitude, latitude, and species, followed by annual attribute columns labeled by year (e.g., 1991, 1992, …).

**year.span**

a short description of the input dataset.This text will appear in the report as data source for the generated figures.


## **generate report:**

```{r st_report_demo, eval=FALSE, message = FALSE, warning = FALSE, results = 'hide'}

# e.g. series to check
generate_report(robj = results_mapping,
                png.text = list(
                  text_top  = "Ring width measurment - ",
                  text_bott = "Source: demo-samples",
                  text_side = "ring width (mm)"
                )
                )
```
#### **arguments of the generate_report() function:**

**robj **

The input for the spatial-temporal report is the output of the CFS_mapping() function, which assigns the class "cfs_mapping" to the resulting object.

**png.text**

A list of text labels for the plot. Supported elements are text_top, text_bott, and text_side.

```{r map_report, include= FALSE, message = FALSE, warning = FALSE, results = 'hide'}


outfile_map <- tempfile(fileext = ".html")

generate_report(robj = results_mapping,
                png.text = list(
                  text_top  = "Ring width measurment - ",
                  text_bott = "Source: demo-samples",
                  text_side = "ring width (mm)"
                ),
                output_file = outfile_map
                )
```

```{r map_inclu, echo = FALSE, results = 'asis'}
# cat('<div style="border: 5px solid #4682B4; border-radius: 8px; padding: 10px; 
#           margin: 15px 0; background-color: #f9f9f9; 
#           max-height: 400px; overflow-y: scroll;">')

cat('
<div style="margin: 20px 0; padding: 10px; 
            box-shadow: 0 2px 8px rgba(0,0,0,0.1); 
            border-radius: 8px;">
')

htmltools::includeHTML(outfile_map)  # embed in vignette
cat('</div>')
```
------------------------------------------------------------------------

