---
title: "Spatio-Temporal Dynamics Report"

date: "`r Sys.Date()`"
output:
  html_document:
  theme: flatly
highlight: tango

params:
  robj: NULL
  animation_fps: !r 1
  data.crs:  "EPSG:4326"
  png.text: !r NULL



---

```{r setup_gif, include=FALSE}
# This chunk sets global options to hide all R code from appearing in the output


library(growthTrendR)
knitr::opts_chunk$set(echo = FALSE)

```

<br>

The following animation illustrates temporal patterns of a key attribute across a geographic region. Values at discrete locations are interpolated using the inverse distance weighting (IDW) method to create continuous yearly surfaces. The animation provides a dynamic view of spatio-temporal trends, highlighting patterns and potential anomalies in the data.
<br>




```{r show_gif, echo=FALSE, results='asis', warning=FALSE, message=FALSE, fig.width=10, fig.height=5, fig.align='left'}


maps <- plot_mapping(mapping_results = params$robj, parms.out = NULL, png.text = params$png.text, animation_fps = params$animation_fps, crs.src = params$data.crs)


if (is.null(maps) || length(maps) == 0) {
  cat("**Warning:** No GIFs found in 'maps' list.\n\n")
} else {
  
  for (gif_name in names(maps)) {
    
    gif_file <- maps[[gif_name]]
    
    # Write to temp file, embed as src
    tmp <- tempfile(fileext = ".gif")
    magick::image_write(gif_file, path = tmp, format = "gif")
    
    cat("### ", gsub("_", " ", gif_name), "\n\n")
    
    cat(sprintf('<div style="text-align: center; margin: 20px 0;">\n'))
    cat(sprintf('  <img src="%s" alt="%s" style="max-width: 100%%; height: auto; border: 1px solid #ddd; border-radius: 4px;">\n',
                tmp, gif_name))
    cat(sprintf('</div>\n\n'))
  }
}
```

